---
title: "Peer 1 Assignment"
output: html_document
---

###Preliminaries

```{r}
library(knitr)
opts_chunk$set(echo = TRUE, results = "hold")

library(data.table)
library(ggplot2)
```

###Loading and preprocessing the data

The data in the *repdata-data-activity.zip* comes from a step counter that gathers personal data at 5 minute intervals throughout the day. The data was collected in October and Novemeber of 2012.

1.Loading the data

```{r}
sdat <- read.csv(unz('repdata-data-activity.zip', 'activity.csv'), header =TRUE)
```

2. Cleaning the Data

```{r}
sdat$date <- as.Date(sdat$date, format = "%Y-%m-%d")
sdat$interval <- as.factor(sdat$interval)
```

###Number of steps taken in a day

1. We can aggregate the number of steps taken per day.


```{r}
daily_steps <- aggregate(steps ~ date, sdat, sum)
colnames(daily_steps) <- c("date", "steps")
head(daily_steps)
```

2. Histogram of Number of steps taken per day.

```{r}
ggplot(daily_steps, aes(x = steps)) + 
  geom_histogram(fill = "red", binwidth = 2000) +
  labs(title = "Histogram of Daily Steps",
       x = "Number of Daily Steps", y = "Number of Days")

```

3. The mean and median of steps taken per day.

```{r}
steps_mean   <- round(mean(daily_steps$steps, na.rm=TRUE))
steps_median <- median(daily_steps$steps, na.rm=TRUE)
```

```{r, echo=FALSE}
steps_mean
steps_median
```

###What is the average daily activity pattern?

1. 

```{r}
dt_steps <- aggregate(x=list(steps=sdat$steps), by = list(interval = sdat$interval), FUN=mean, na.rm=TRUE)

head(dt_steps)
dt_steps$interval <- 
        as.integer(levels(dt_steps$interval)[dt_steps$interval])
colnames(dt_steps) <- c("interval", "steps")
```

We can now plot the number of steps taken in a 5 minute interval.

```{r}
ggplot(dt_steps, aes(x=interval, y=steps)) +   
        geom_line(color="blue", size=1) +  
        labs(title="Average Daily Activity Pattern", x="Time x 5 [min]", y="Number of steps")
```

2. The maximum number of steps can be readily calculated.

```{r}
max_steps <- dt_steps[which.max(dt_steps$steps),]
```
```{r,echo=F}
echo=FALSE
max_steps$steps
```

##Imputting missing values

1. The number of missing *NA* values is simply:

```{r}
na_steps <- sum(is.na(sdat$steps))
```

```{r, echo=FALSE}
na_steps
```

2. Fill the data with the **fill_na** function which utilizes **mapply**. A filled set is returned that replaces all *NA* values with the average value for that interval calculated in *dt_steps*.

```{r}
fill_na <- function(steps, interval){
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (dt_steps[dt_steps$interval==interval, "steps"])
  return(filled)
}
```

3. A new data set *sData* has *NA* replaced with the averages.

```{r}
sData <- sdat
sData$steps <-mapply(fill_na, sData$steps, sData$interval)
```

```{r}
head(sData)
```

4. Plot new histogram with filled-in values.

```{r}
sData_sum <- aggregate(steps ~ date, sData, sum)
ggplot(sData_sum, aes(x = steps)) + 
  geom_histogram(fill = "red", binwidth = 2000) +
  labs(title = "Histogram of Daily Steps",
       x = "Number of Daily Steps", y = "Number of Days")

sD <- tapply(sData$steps, sData$date, FUN=sum)
mean(sD)
median(sD)
```

##Weekdays vs. Weekends

1. We now write a function that determines wheter or not a given day is a weekday or weekend.

```{r}
whatday <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Sunday", "Saturday"))
        return("weekend")
    else 
        return("weekday")
}

```

2. Append either weekend or weekday to the data frame.

```{r}
sData$date <- as.Date(sData$date)
sData$day <- sapply(sData$date, FUN=whatday)
```

3. Plot two histograms with a filter on weekday or weekend.

```{r}
averages <- aggregate(steps ~ interval + day, data=sData, mean)
averages$interval <- 
        as.integer(levels(averages$interval)[averages$interval])
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("Time x 5 [min]") + ylab("Number of steps")
```